---
title: "Using clustur"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using-clustur}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# mothur
mothur is an open-sourced software used for the analysis of ecological data. It has a strong presence in the use of analyzing 16S rRNA gene sequences.
mothur itself has a plethora of features, such as, clustering based on genetic similarity, sequence alignment, taxomic assignment, alpha and beta diversity,
and phylotype analysis. There have been numerous publications using this software by individuals and intuitions alike. We plan to take advantage of the power of
mothur, and isolate the clustering feature. We believe this will allow scientist to work with a familiar tool and give rise for the creation of more pipelines
utilizing this package.


# Clustur
clustur was developed for two main reasons: to allow developers to easily create pipelines with mothur code in R, and to give users a powerful
clustering algorithm from mothur without downloading and building it. Because mothur is built using rcpp, it benefits from the speed of c++
and the usability of R.


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Opticluster)
```
